---
title: "Garnett"
---

# do this in the terminal using "sudo R"
```{r}
# First install Bioconductor and Monocle
if (!requireNamespace("BiocManager"))
    install.packages("BiocManager")

BiocManager::install()
BiocManager::install(c("monocle"))

# Next install a few more dependencies
BiocManager::install(c('DelayedArray', 'DelayedMatrixStats', 'org.Hs.eg.db', 'org.Mm.eg.db'))
```



# install garnett "sudo R"
```{r}
install.packages("devtools")
devtools::install_github("cole-trapnell-lab/garnett")
```



# load garnett
```{r}
library(garnett)
```

# load monocle
```{r}
library(monocle)
```

# install hdf5r using "sudo R"
```{r}
devtools::install_github("hhoeflin/hdf5r")
```

# load hdf5r
```{r}
library(hdf5r)
```

# install Seurat using "sudo R"
```{r}
install.packages('Seurat')
```

# load dplyr
```{r}
library(dplyr)
```

# load Seurat
```{r}
library(Seurat)
```

# load monocle
```{r}
library(monocle)
```


# check versions of things you loaded
```{r}
sessionInfo()
```


# extract matrix of raw counts for 
```{r}
MatrixForGarnett <- as.matrix(GetAssayData(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP, slot = "counts"))
                              #[, WhichCells(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR, ident = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25", "26", "27"))])
#https://satijalab.org/seurat/v3.0/interaction_vignette.html
#https://github.com/satijalab/seurat/issues/2362

```

```{r}
MatrixForGarnett_sparse <- Matrix(MatrixForGarnett, sparse = TRUE)
```

```{r}
write.table(as.matrix(MatrixForGarnett_sparse), file="MatrixForGarnett_sparse.csv", quote=FALSE, sep = ",", row.names=TRUE)
```


```{r}
write.table(as.matrix(MatrixForGarnett_sparse), file="MatrixForGarnett_sparse.txt", quote=FALSE, sep = "\t", row.names=TRUE)
```



# write table of gene names that are present
```{r}
write.table(as.data.frame(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP@assays[["RNA"]]@data@Dimnames[[1]]),
          file="GeneNamesInGarnettMatrix.txt", quote=FALSE, sep = "\t", row.names=FALSE)
```


# generate featureData
```{r}
write.table(as.data.frame(row.names(MatrixForGarnett)),
          file="FeatureData.txt", quote=FALSE, sep = "\t", row.names=FALSE)
```

# generate phenoData
```{r}
write.table(colnames(MatrixForGarnett),
          file="PhenoData.tsv", quote=FALSE, sep = '\t', row.names=FALSE)
```





# load the data
```{r}
mat <- read.delim("MatrixForGarnett_sparse.txt", row.names = 1)
pdata <- read.delim("PhenoData_Labeled_06.17.20.txt", row.names = 1)
fdata <- read.delim("FeatureData_Correct.txt", row.names = 1)
```


# check data #1
```{r}
head(mat)
```


# check data #2
```{r}
head(pdata)
```

# check data #3
```{r}
head(fdata)
```

# create CellDataSet
```{r}
pd <- new("AnnotatedDataFrame", data = pdata)
fd <- new("AnnotatedDataFrame", data = fdata)
cds <- newCellDataSet(as.matrix(mat),
    phenoData = pd, featureData = fd)
```



```{r}
# generate size factors for normalization later
cds <- estimateSizeFactors(cds)
```

## Load and Check Marker file ##

# load required argument for a Bioconductor AnnotationDb-class package used for converting gene IDs
```{r}
library(org.Hs.eg.db)
```


# load marker file
```{r}
marker_file <- read.delim("MarkerFile_Pancreas_Garnett_16.txt")
```

# check if markers are good ones
```{r}
marker_check <- check_markers(cds, "MarkerFile_Pancreas_Garnett_16.txt",
                              db=org.Hs.eg.db,
                              cds_gene_id_type = "SYMBOL",
                              marker_file_gene_id_type = "SYMBOL")
```

```{r}
pdf('marker_check_v12.pdf')
plot_markers(marker_check)
dev.off()
```

# TRAIN CLASSIFIER
```{r}
#https://github.com/cole-trapnell-lab/monocle3/issues/272

library(org.Hs.eg.db)
set.seed(260)


classifier_500 <-train_cell_classifier(cds = cds,
                                         marker_file = "MarkerFile_Pancreas_Garnett_16.txt",
                                         db=org.Hs.eg.db,
                                         cds_gene_id_type = "SYMBOL",
                                         num_unknown = 500,
                                         marker_file_gene_id_type = "SYMBOL", cores = 35)
```


```{r}
head(pData(cds))
```

# Viewing the classification genes when 100 cells are used for classifying

```{r}
feature_genes_500 <- get_feature_genes(classifier_500,
                                   node = "root",
                                   db = org.Hs.eg.db, 
                                   convert_ids = TRUE)
head(feature_genes_500)
```



## chose classifier that worked with 150 cells because the gene list seemed most specific

## classify cells!

```{r}
library(org.Hs.eg.db)
cds_500 <- classify_cells(cds, classifier_500,
                           db = org.Hs.eg.db,
                           cluster_extend = TRUE,
                           cds_gene_id_type = "SYMBOL")
```


```{r}
head(pData(cds_500))
```

# write a table
```{r}
write.table(as.data.frame(pData(cds_500)),
          file="Barcodes_GarnettLabels_cds_500_6.24.20.txt", quote=FALSE, sep = "\t", row.names=TRUE)
```

# read in the Garnett labels
```{r}
cell_type <- read.delim("cell_type_06.24.20.txt", row.names = 1)
```


# read in the Garnett labels
```{r}
cluster_ext_type <- read.delim("ext_cell_type_06.24.20.txt", row.names = 1)
```



```{r}
table(pData(cds_500)$cell_type)
```

```{r}
table(pData(cds_500)$cluster_ext_type)
```

# install UMAP
```{r}
library(reticulate)
use_condaenv(condaenv="Renv", conda="/mnt/data0/apps/anaconda/Anaconda2-5.2/bin/conda")
library(Seurat)
```


```{r}
reticulate::py_install(packages = 'umap-learn')
```

```{r}
All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP$cell_type <- cell_type
```


```{r}
All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP$cluster_ext_type <- cluster_ext_type
```


```{r}
All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap <- RunUMAP(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP, dims = 1:26)
```

```{r}
cell_type.info <- data.frame(cell_type = cell_type, row.names= colnames(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap))
```

```{r}
All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap <- AddMetaData(object = All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap, metadata = cell_type.info)
```


```{r}
pdf('GarnetLabels_UMAP_cell_type_final_06.24.20.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap, group.by = "cell_type", reduction = "umap", label = FALSE, cols= c("#e30800", "#f56505", "#dec400", "#006630", "#0223c7","#5b02c7", "#00b0e6", "#c40080", "#02f00a", "#7d3301", "#000000"))
dev.off()
```


```{r}
cluster_ext_type.info <- data.frame(cluster_ext_type = cluster_ext_type, row.names= colnames(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap))
```

```{r}
All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap <- AddMetaData(object = All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap, metadata = cluster_ext_type.info)
```


```{r}
pdf('GarnetLabels_UMAP_ext_cell_type_final_06.24.20.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#e30800", "#f56505", "#dec400", "#006630", "#0223c7","#5b02c7", "#00b0e6", "#c40080", "#02f00a", "#7d3301", "#000000"))
dev.off()
```


###### didn't do the below ######


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_ALPHA_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#fac720", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```

```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_BETA_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#fac720", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```

```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_DELTA_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#fac720", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```

```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_DUCTAL_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#fac720","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_ENDOTHELIAL_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#fac720", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_EPSILON_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#fac720", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_IMMUNE_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#fac720", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_MACRO_MONO_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#fac720", "#000000", "#000000", "#000000")) 
dev.off()
```


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_PP_Gamma_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#fac720", "#000000", "#000000")) 
dev.off()
```


```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_Stellates_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#fac720", "#000000")) 
dev.off()
```



```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_Unknown_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#000000", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#fac720")) 
dev.off()
```





```{r}
pdf('GarnetLabels_UMAP_cluster_ext_type_ACINAR_750.pdf')
DimPlot(All_Islet_DoubletRemoved_Norm_VariableGenes_Scaled_PCA_Jack_Cluster_UMAP_GROUP_DONOR_umap, group.by = "cluster_ext_type", reduction = "umap", label = FALSE, cols= c("#fac720", "#000000", "#000000", "#000000", "#000000","#000000", "#000000", "#000000", "#000000", "#000000", "#000000", "#000000")) 
dev.off()
```